---
title: "Neotropical Metabarcoding Metadata"
author: "Frédéric Mahé"
date: '`r format(Sys.time(), "%d %B %Y")`'

output:
  rmarkdown::html_document:
    theme: lumen
    toc: yes
    toc_float: TRUE
    keep_md: yes
    # code_folding: hide
---

```{r setup, include=FALSE}
rm(list = ls()) # remove all objects before starting
knitr::opts_chunk$set(echo = TRUE)
```


#### load required packages

```{r packages, message=FALSE}
library(here)
library(tidyverse)
```

# preliminary steps

### variables and functions

```{r}
eukbank_metadata_file <- "eukbank_18SV4_asv.metadata"
coordinates_file <- "sample_coordinates.tsv"
project_name_in_eukbank <- "Neotropical_lowland_forests"
list_of_read_files <- "list_of_read_files.tsv"
## output <- str_replace(input, "..xlsx", "_reduced_taxonomy.tsv")
## annotated_table <- str_replace(output, ".tsv", ".xlsx")
## eukaryotic_rarefied_file <- str_replace(input, "..xlsx", "_eukaryotic_rarefied.tsv")
## parasitic_rarefied_file <- str_replace(input, "..xlsx", "_parasitic_rarefied.tsv")
```

```{r}
load_list_of_read_files <- function(filename){
    here::here("data", filename) %>%
        read_tsv(show_col_types = FALSE) %>%
        mutate(sample = str_remove(string = Submitted_files,
                                    pattern = "_lib.*"),
               sample = str_remove(string = sample,
                                    pattern = "NG-[0-9]+_")) %>%
        filter(str_detect(string = Submitted_files,
                          pattern = "_1_2\\.",
                          negate = TRUE))
}


load_per_sample_coordinates <- function(filename){
    here::here("data", filename) %>%
        read_tsv(show_col_types = FALSE) %>%
        mutate(northing = as.character(northing),
               sample_description = str_glue(
                   "Sampling position of {sample}",
                   "in UTM coordinates are zone: {zone},",
                   "northing: {northing}, easting: {easting}.",
                   .sep = " ")
               )
}


load_raw_metadata <- function(filename){
    here::here("data", filename) %>%
        read_tsv(show_col_types = FALSE) %>%
        filter(project == project_name_in_eukbank) %>%
        select(-c(project, notus, nreads, size_fraction_lower_threshold,
                  size_fraction_upper_threshold))
}
```


### get a list of available read files (sff and fastq)

```{r}
load_list_of_read_files(list_of_read_files) -> read_file_names
```


Some samples were sequenced twice (Roche 454 and Illumina
MiSeq). After extracting sample names from read file names, we expect
to find 154 unique sample names.

```{r}
read_file_names %>%
    distinct(sample) %>%
    pull() %>%
    length() -> n_unique_samples
```

There are `r prettyNum(n_unique_samples, scientific=FALSE,
big.mark=",")` unique sample names.


### get per-sample coordinates

Using the Universal Transverse Mercator (UTM) coordinate system:

```{r}
load_per_sample_coordinates(coordinates_file) -> sample_cordinates
```

# load metadata

```{r}
load_raw_metadata(eukbank_metadata_file) -> metadata
```

## check

Are sample lists equal?

```{r}
setequal(x = metadata %>%
             distinct(sample),
         y = read_file_names %>%
             distinct(sample))
```

Yes, the metadata table and the list of available read files contain
the same lists of sample names.


## create the 'sample description' field

```{r}
sample_cordinates %>%
    pull(sample_description)

## create for double samples!
metadata %>%
    select(sample) %>%
    filter(str_detect(string = sample, pattern = "_")) %>%
    mutate(sample_pair = sample) %>%
    separate_rows(sample, sep = "_") %>%
    left_join(sample_cordinates %>%
              select(sample, sample_description),
              by = "sample") %>%
    select(-sample) %>%
    group_by(sample_pair) %>%
    mutate(position = row_number(sample_pair)) %>%
    ungroup() %>%
    pivot_wider(names_from = position, values_from = sample_description)

## almost there!
```


***

```{r}
sessionInfo()
rm(list = ls())
```
